---
title: "IPHC data - halibut"
author: "Alberto Rovellini"
date: "01/21/2022"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(rbgm)
library(sf)
library(viridis)
library(maps)
library(mapdata)
library(data.table)
```

```{r}
select <- dplyr::select
```

Explore the IPHC longline data, to get some distributions of Pacific Halibut. These will be compared with the distributions that we got from bottom trawl surveys for dogfish.  

Need to look into what this RPN business is.
```{r}
dat <- read.csv('../data/IPHC/fiss_cleaned_09222021.csv')

atlantis_bgm <- read_bgm('../data/GOA_WGS84_V4_final.bgm')
atlantis_box <- atlantis_bgm %>% box_sf()
# utilities
atlantis_crs <- atlantis_bgm$extra$projection
atlantis_bbox <- atlantis_box %>% st_bbox()
```

Check how many data points there are for each species, and which years we have surveys for.
```{r}
dat %>% select(SPECIES_COMMON_NAME) %>% group_by(SPECIES_COMMON_NAME) %>% tally() %>% arrange(desc(n))
dat %>% select(SURVEY_YEAR) %>% distinct() %>% pull() %>% sort()
```

Extract halibut.

The data reports the weight of legal and sublegal halibut per set. It also reports the number of extrapolated effective hooks (the hooks that were fishing effectively). So we can get a total halibut catch (legal and sublegal) and then standardize it by the number of effective hooks in the set. 

```{r}
#dat %>% select(SPECIES_COMMON_NAME) %>% distinct() %>% arrange(SPECIES_COMMON_NAME) %>%  pull()

halibut <- dat %>% 
  filter(SPECIES_COMMON_NAME=='Pacific Halibut' & IPHC_REG %in% c('4A','3B','3A','2C','2B')) %>%
  select(FISHING_EVENT_ID,SURVEY_YEAR,IPHC_REG,START_LAT,START_LON,AVG_DEPTH_M,EX_EFF_HOOKS,LGL_HAL_WT,SUBLGL_HAL_WT) %>%
  mutate(HAL_WT = LGL_HAL_WT + SUBLGL_HAL_WT,
         HAL_WT_STD = HAL_WT / EX_EFF_HOOKS)
```

View in space.
```{r}
coast <- maps::map("worldHires", c("USA","Canada"), plot = FALSE, fill = TRUE)
coast_sf <- coast %>% st_as_sf() #%>% st_transform(crs = atlantis_crs)

halibut_sf <- halibut %>% st_as_sf(coords=c('START_LON','START_LAT'),crs=4326)
halibut_bbox <- halibut_sf %>% st_bbox()

halibut_sf %>% ggplot()+
  geom_sf(aes(color=HAL_WT_STD))+
  scale_color_viridis()+
  geom_sf(data=coast_sf)+
  coord_sf(xlim=c(halibut_bbox$xmin,halibut_bbox$xmax), ylim=c(halibut_bbox$ymin,halibut_bbox$ymax))+
  facet_wrap(~SURVEY_YEAR)+
  theme_bw()
```

Let's get the fishing events in these Regulatory Areas that did not catch any halibuts.
```{r}
these_events <- dat %>% filter(IPHC_REG %in% c('4A','3B','3A','2C','2B')) %>%
  select(FISHING_EVENT_ID) %>% 
  distinct() %>% 
  pull()

halibut_events <- halibut %>%  select(FISHING_EVENT_ID) %>%  distinct() %>% pull()

# build a dataframe with events that had zero catch for halibut
no_halibut_events <- setdiff(these_events,halibut_events)

no_halibut_frame <- dat %>% 
  filter(FISHING_EVENT_ID %in% no_halibut_events & IPHC_REG %in% c('4A','3B','3A','2C','2B')) %>% 
  select(FISHING_EVENT_ID,SURVEY_YEAR,IPHC_REG,START_LAT,START_LON,AVG_DEPTH_M,EX_EFF_HOOKS,LGL_HAL_WT,SUBLGL_HAL_WT) %>%
  mutate(HAL_WT = LGL_HAL_WT + SUBLGL_HAL_WT,
         HAL_WT_STD = HAL_WT / EX_EFF_HOOKS) %>%
  distinct() 

halibut_all <- rbind(halibut,no_halibut_frame) # slightly longer than the sum of events, meaning that some events are duplicated. Explore below

halibut_all <- halibut_all %>% distinct() # this gets rid of most duplicate entries

# dups <- halibut_all %>% group_by(FISHING_EVENT_ID) %>% tally() %>% filter(n>1) %>% select(FISHING_EVENT_ID) %>% pull()
# dups_halibut <- halibut_all %>% filter(FISHING_EVENT_ID %in% dups) %>% distinct() 
```

View.
```{r}
halibut_all_sf <- halibut_all %>% st_as_sf(coords=c('START_LON','START_LAT'),crs=4326)
halibut_all_bbox <- halibut_all_sf %>% st_bbox()

halibut_all_sf %>% ggplot()+
  geom_sf(aes(color=HAL_WT_STD))+
  scale_color_viridis()+
  geom_sf(data=coast_sf)+
  coord_sf(xlim=c(halibut_bbox$xmin,halibut_bbox$xmax), ylim=c(halibut_bbox$ymin,halibut_bbox$ymax))+
  facet_wrap(~SURVEY_YEAR)+
  theme_bw()
```
This is a big dataset with a lot of points. May be worth trying with sdmTMB, the issue is getting a CPUE from the hook counts.

Prepare this for the sdmTMB code.
```{r}
halibut_all <- halibut_all %>% select(SURVEY_YEAR,FISHING_EVENT_ID,START_LAT,START_LON,AVG_DEPTH_M,HAL_WT_STD) %>% 
  set_names(c(
  "year",
  "hauljoin",
  "lat",
  "lon",
  "depth",
  "cpue"))

saveRDS(halibut_all,'../data/IPHC/halibut.RDS')
```
